R Notebook

### 2023-11-16 02_paper/02_study/02_fitbit paper/analysis_270423.Rmd
update fitbit paper

main <- 
  main %>% 
  dplyr::select("time", "heart_rate", "ID", "time_span") %>% 
  distinct() %>% 
  group_by(ID) %>% 
  dplyr::summarize(mean_heart = mean(heart_rate),
            sd_heart = sd(heart_rate)) %>%
  ungroup() %>% 
  right_join(x = .,
             y = main,
             by = "ID") %>% 
  mutate(heart_rate_std = (heart_rate - mean_heart)/
           sd_heart) %>% 
  filter(!(time_span == "teaching" & time > 600))

@@

```{r fig.width = 10,results='hide',fig.keep='all'}
loess_plot_overall <-
  main %>% 
  filter(time_span == "overall") %>%
  dplyr::select("ID","time_span", "time","heart_rate_std") %>% 
  distinct() %>% 
  ggplot(mapping = aes(x = time,
                       y = heart_rate_std
                       )
         ) +
  # geom_point(alpha = 0.05) + 
  geom_smooth(se = FALSE,
              size = 0.5,
              colour = "black") +
  scale_color_viridis_d(option = "A",
                        end = 0.8) +
  scale_linetype_manual(values = c(1, 6)) +
  scale_x_continuous(expand = c(0, NA),
                     breaks = seq(0, 9000, 600)) +
  scale_y_continuous(breaks = seq(-1, 1, 0.5)) + 
  labs(x = "Time (in Seconds)",
       y = "") +
  ggtitle("Overall Course of Average Heartrate") +
  theme_apa() + 
  theme(axis.title.y = element_blank())

@@

ki_plot_phases <-
  main %>% 
  filter(time_span != "overall") %>%
  dplyr::select("ID","time_span", "time","heart_rate_std") %>% 
  mutate(time_span = factor(time_span,
                            levels = c("preparation","teaching","post","interview","end"),
                            labels = c("(1) Pre-teaching phase","(2) Teaching phase","(3) Post-teaching phase","(4) Interview phase","(5) End phase"))) %>% 
  distinct() %>%
  group_by(time_span) %>% 
  summarise(z_mean = mean(heart_rate_std),
            z_sd = sd(heart_rate_std),
            z_t = qt(p = 0.01 / 2,      # 0.01 KI
               df = n() - 1,
               lower.tail = F
               ),
            z_error = z_sd / sqrt(n()),
            lower = z_mean - z_t * z_error,
            upper = z_mean + z_t * z_error,
            ) %>% 
  ungroup() %>% 
  ggplot(mapping = aes(x = factor(0),
                       y = z_mean)) + 
  geom_point() +
  geom_errorbar(mapping = aes(ymin = lower,
                              ymax = upper),
                width = 0.2) +
  geom_hline(yintercept = 0,
             linetype = "dashed") +
  labs(subtitle = "1% Confidence Intervall",
       x = "") +
  facet_grid(cols = vars(time_span)) +
  theme_apa() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank())

@@

lm_plot <-
  main %>%
  filter(time_span != "overall") %>%
  dplyr::select("ID","time_span", "time","heart_rate_std") %>%
  mutate(time_span = factor(time_span,
                            levels = c("preparation","teaching","post","interview","end"),
                            labels = c("(1) Pre Teaching\nPhase","(2) Teaching\nPhase","(3) Post Teaching\nPhase","(4) Interview\nPhase","(5) End\nPhase")
                            ),
         time = time / 60
         ) %>%
  distinct() %>%
  ggplot(mapping = aes(x = time,
                       y = heart_rate_std
                       )
         ) +
  # geom_point(alpha = 0.05) +
  geom_smooth(method = "lm",
              se = T,
              size = 0.5,
              colour = "black",
              level = 0.99) +
  scale_color_viridis_d(option = "A",
                        end = 0.8) +
  #scale_linetype_manual(values = c(1, 6)) +
  scale_x_continuous(breaks = seq(0, 10, 2)) +
  #scale_y_continuous(breaks = seq(0, 160, 5)) +
  labs(x = "Time (in Minutes)",
       y = "Average Standardized Heart Rate",
       caption = expression(paste(italic("Note: "),
                                  "The shadow around the line represents the 99% confidence interval."
                                  )
                            )
       ) +
  theme_apa() +
  theme(plot.subtitle = element_text(face = "italic"),
        plot.caption = element_text(hjust = 0)) +
  facet_grid(cols = vars(time_span),
             scales = "free_x")
ggsave(filename = "plots_publication/lm_plot.png",
       plot = lm_plot,
       width = 20,
       height = 12,
       units = "cm")

@@

```{r}
main %>% 
  filter(time_span == "preparation") %>% 
  group_by(ID) %>% 
  transmute(heart_rate_m = mean(heart_rate_std),
            confidence_factor_m = mean(confidence_factor,
                                      na.rm = TRUE),
            disruption_factor_m = mean(disruption_factor,
                                       na.rm = TRUE),
            teaching_experience = teaching_experience,
            gender = gender
  ) %>%
  distinct() %>% 
  ungroup() %>% 
  transmute(heart_rate_m = (heart_rate_m - mean(.$heart_rate_m))/sd(.$heart_rate_m),
         confidence_factor_m = (confidence_factor_m - mean(.$confidence_factor_m))/sd(.$confidence_factor_m),
         disruption_factor_m = (disruption_factor_m - mean(.$disruption_factor_m))/sd(.$disruption_factor_m),
         gender = if_else(gender == "male",
                           true = 0,
                           false = 1),
          teaching_experience = (teaching_experience - mean(.$teaching_experience))/sd(.$teaching_experience)
         ) %>%
  pcor()  -> pcor_pre

@@

@@

# left_join(x = df_heart_rate,
#           y = df_heart_rate_std,
#           by = c("ID","time")) %>% 
bind_rows(df_heart_rate,
          df_heart_rate_std
          ) %>% 
  ggplot(mapping = aes(x = time,
                       y = heart_rate
                       )
         ) +
  # geom_point(alpha = 0.05) + 
  geom_smooth(se = T,
              size = 0.5,
              colour = "black",
              level = 0.99) +
  scale_color_viridis_d(option = "A",
                        end = 0.8) +
  scale_linetype_manual(values = c(1, 6)) +
  # scale_x_continuous(expand = c(0, NA),
  #                    breaks = seq(0, 120, 10)) +
  scale_x_continuous(expand = c(0.01,0)) + 
  labs(x = "Time (in minutes)",
       #y = "Heart Rate (in Beats per Minute)",
       y = "",
       # title = "Figure #",
       # subtitle = "Overall Course of Average Heartrate",
       # caption = expression(paste(italic("Note: "),
       #                            "The shadow around the line represents the 99% confidence interval. 
       #                            We used the ggplot2 package (v3.3.3; Wickham, 2016) to calculate the moving average of the course."))
       ) +
  #ggtitle("Overall Course of Average Heartrate") +
  theme_apa() + 
  # theme(plot.subtitle = element_text(face = "italic"),
  #       plot.caption = element_text(hjust = 0),
  #       axis.line = element_line(colour = "black"),
  #       panel.grid.major = element_blank(),
  #       panel.grid.minor = element_blank(),
  #       panel.border = element_blank(),
  #       panel.background = element_blank()) +
  facet_wrap(~ type,
             ncol = 2,
             scales = "free_y")

@@

```{r}
  main %>% 
  filter(time_span != "overall") %>%
  dplyr::select("ID","time_span", "time","heart_rate_std") %>% 
  mutate(time_span = factor(time_span,
                            levels = c("preparation","teaching","post","interview","end"),
                            labels = c("(1) Pre-teaching\nphase","(2) Teaching\nphase","(3) Post-teaching\nphase","(4) Interview\nphase","(5) End\nphase"))) %>% 
  distinct() %>%
  group_by(time_span) %>% 
  summarise(z_mean = mean(heart_rate_std),
            z_sd = sd(heart_rate_std),
            z_t = qt(p = 0.01 / 2,      # 0.01 KI
               df = n() - 1,
               lower.tail = F
               ),
            z_error = z_sd / sqrt(n()),
            lower = z_mean - z_t * z_error,
            upper = z_mean + z_t * z_error,
            ) %>% 
  ungroup() %>% 
  ggplot(mapping = aes(x = factor(0),
                       y = z_mean)) + 
  geom_point() +
  geom_errorbar(mapping = aes(ymin = lower,
                              ymax = upper),
                width = 0.2) +
  geom_hline(yintercept = 0,
             linetype = "dashed") +
  labs(x = "",
       y = "Standardized mean HR",
       caption = expression(paste(italic("Note: "), 
                                  "The dotted line represents the null effect; bars represent the 1% confidence interval. "
                                  )
                            )
       ) +
  facet_grid(cols = vars(time_span)) +
  theme_apa() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.caption = element_text(hjust = 0))

@@

@@

```{r echo = FALSE, results='asis'}
df_multi <- 
  main %>% 
  select("ID", "disruption_factor", "confidence_factor", "teaching_experience") %>% 
  distinct() %>% 
  group_by(ID) %>% 
  summarize(disruption = mean(disruption_factor, na.rm = T),
            confidence = mean(confidence_factor, na.rm = T),
            teaching_experience) %>% 
  ungroup() %>% 
  mutate(id = as.character(ID)) %>% 
  distinct() %>% 
  right_join(x = .,
            y = fix_models,
            by = "id") %>% 
  mutate(posneg = if_else(condition = slope < 0,
                          true = "negative",
                          false = "positive"))
fit_teach_1 <-
  df_multi %>% 
  filter(Phase == "Teaching Phase") %>% 
  split(.$posneg) %>% 
  map(~ lm(slope ~ disruption,
           data = .))

>>>>

main <- 
  main %>% 
  dplyr::select("time", "heart_rate", "ID", "time_span") %>% 
  distinct() %>% 
  group_by(ID) %>% 
  dplyr::summarize(mean_heart = mean(heart_rate),
            sd_heart = sd(heart_rate)) %>%
  ungroup() %>% 
  right_join(x = .,
             y = main,
             by = "ID") %>% 
  filter(!(time_span == "teaching" & time > 600)) %>%
  mutate(heart_rate_std = (heart_rate - mean_heart)/
           sd_heart)

# Descriptives
```{r descriptive statistics}

# Mean, SD and range for overall (unstandardized)
descrip_hr_overall <- main %>%
  filter(time_span %in% c("overall")) %>% 
  summarise(N = n_distinct(ID),
            "M heart rate in bpm" = round(mean(heart_rate),
                                     digits = 2),
            "SD heart rate in bpm" = round(sd(heart_rate),
                                      digits = 2),
            "Min heart rate in bpm" = min(heart_rate),
            "Max heart rate in bpm" = max(heart_rate)
  )

# Mean, SD and range for overall (standardized)
descrip_hr_overall_std <- main %>%
  filter(time_span %in% c("overall")) %>% 
  summarise(N = n_distinct(ID),
            "M std heart rate in bpm" = round(mean(heart_rate_std),
                                              digits = 2),
            "SD std heart rate in bpm" = round(sd(heart_rate_std),
                                               digits = 2),
            "Min std heart rate in bpm" = round(min(heart_rate_std),
                                                digits = 2),
            "Max heart rate in bpm" = round(max(heart_rate_std),
                                            digits = 2)
            )

# Mean, SD and range for individual phases (unstandardized)
descrip_hr_phases <- main %>% 
  filter(!time_span %in% c("overall")) %>% 
  group_by(time_span) %>% 
  summarise(N = n_distinct(ID),
            "M heart rate in bpm" = round(mean(heart_rate),
                                     digits = 2),
            "SD heart rate in bpm" = round(sd(heart_rate),
                                      digits = 2),
            "Min heart rate in bpm" = min(heart_rate),
            "Max heart rate in bpm" = max(heart_rate)
  )

# Mean, SD and range for individual phases (standardized)
descrip_hr_phases_std <- main %>% 
  filter(!time_span %in% c("overall")) %>% 
  group_by(time_span) %>% 
  summarise(N = n_distinct(ID),
            "M heart rate in bpm" = round(mean(heart_rate_std),
                                     digits = 2),
            "SD heart rate in bpm" = round(sd(heart_rate_std),
                                      digits = 2),
            "Min heart rate in bpm" = round(min(heart_rate_std),
                                            digits = 2),
            "Max heart rate in bpm" = round(max(heart_rate_std),
                                            digits = 2)
  )

# # format and insert table in manuscript
# descrip_hr_phases %>%
# papaja::apa_table(
#   caption = "Mean, SD and range for teachers' heart rate in individual phases",
#   # note = "Write Note here",
#   escape = TRUE, # if TRUE special Latex characters are escaped; if this is turned to F captions cannot be rendered. I don't know why...
#   placement = "h", # position of table in page:  exact location (h), at the top (t), bottom (b)
#   font_size = "tiny" # options are tiny, scriptsize, footnotesize, small, normalsize (default), large, Large, LARGE, huge, Huge
#   )

# Mean, SD and range for disruption and confidence factor
rating_table <- 
  df_rating %>% 
  filter(!is.na(confidence_factor),
         !is.na(disruption_factor)) %>% 
   summarise(N = n_distinct(ID), 
            "M disruption factor" = round(mean(disruption_factor),
                                     digits = 2),
            "SD disruption factor" = round(sd(disruption_factor),
                                      digits = 2),
            "Min disruption factor" = min(disruption_factor),
            "Max disruption factor" = max(disruption_factor),
            "M confidence factor" = round(mean(confidence_factor),
                                     digits = 2),
            "SD confidence factor" = round(sd(confidence_factor),
                                      digits = 2),
            "Min confidence factor" = min(confidence_factor),
            "Max confidence factor" = max(confidence_factor)
            )

@@

```{r fig.width = 10,results='hide',fig.keep='all'}
loess_plot_overall <-
  main %>% 
  filter(time_span == "overall") %>%
  dplyr::select("ID","time_span", "time","heart_rate_std") %>% 
  distinct() %>% 
  ggplot(mapping = aes(x = time,
                       y = heart_rate_std
                       )
         ) +
  # geom_point(alpha = 0.05) + 
  geom_smooth(se = FALSE,
              size = 0.5,
              colour = "black") +
  scale_color_viridis_d(option = "A",
                        end = 0.8) +
  scale_linetype_manual(values = c(1, 6)) +
  scale_x_continuous(expand = c(0, NA),
                     breaks = seq(0, 9000, 600)) +
  scale_y_continuous(breaks = seq(-1, 1, 0.5)) + 
  labs(x = "Time (in Seconds)",
       y = "") +
  ggtitle("Overall Course of Mean Heart Rate") +
  theme_apa() + 
  theme(axis.title.y = element_blank())

@@

ki_plot_phases <-
  main %>% 
  filter(time_span != "overall") %>%
  dplyr::select("ID","time_span", "time","heart_rate_std") %>% 
  mutate(time_span = factor(time_span,
                            levels = c("preparation","teaching","post","interview","end"),
                            labels = c("(1) Pre-teaching Interval","(2) Teaching Interval","(3) Post-teaching Interval","(4) Interview Interval","(5) End Interval"))) %>% 
  distinct() %>%
  group_by(time_span) %>% 
  summarise(z_mean = mean(heart_rate_std),
            z_sd = sd(heart_rate_std),
            z_t = qt(p = 0.01 / 2,      # 0.01 KI
               df = n() - 1,
               lower.tail = F
               ),
            z_error = z_sd / sqrt(n()),
            lower = z_mean - z_t * z_error,
            upper = z_mean + z_t * z_error,
            ) %>% 
  ungroup() %>% 
  ggplot(mapping = aes(x = factor(0),
                       y = z_mean)) + 
  geom_point() +
  geom_errorbar(mapping = aes(ymin = lower,
                              ymax = upper),
                width = 0.2) +
  geom_hline(yintercept = 0,
             linetype = "dashed") +
  labs(subtitle = "1% Confidence Intervall",
       x = "") +
  facet_grid(cols = vars(time_span)) +
  theme_apa() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank())

@@

lm_plot <-
  main %>%
  filter(time_span != "overall") %>%
  dplyr::select("ID","time_span", "time","heart_rate_std") %>%
  mutate(time_span = factor(time_span,
                            levels = c("preparation","teaching","post","interview","end"),
                            labels = c("(1) Pre-Teaching\nInteval","(2) Teaching\nInterval","(3) Post-Teaching\nInterval","(4) Interview\nInterval","(5) End\nInterval")
                            ),
         time = time / 60
         ) %>%
  distinct() %>%
  ggplot(mapping = aes(x = time,
                       y = heart_rate_std
                       )
         ) +
  # geom_point(alpha = 0.05) +
  geom_smooth(method = "lm",
              se = T,
              size = 0.5,
              colour = "black",
              level = 0.99) +
  scale_color_viridis_d(option = "A",
                        end = 0.8) +
  #scale_linetype_manual(values = c(1, 6)) +
  scale_x_continuous(breaks = seq(0, 10, 2)) +
  scale_y_continuous(breaks = seq(from = -1, to = 1, by = 0.5)) +
  labs(x = "Time (in Minutes)",
       y = "Standardized Mean Heart Rate",
       caption = expression(paste(italic("Note: "),
                                  "The shadow around the line represents the 99% confidence interval."
                                  )
                            )
       ) +
  theme_apa() +
  theme(plot.subtitle = element_text(face = "italic"),
        plot.caption = element_text(hjust = 0)) +
  facet_grid(cols = vars(time_span),
             scales = "free_x")

ggsave(filename = "plots_publication/lm_plot.png",
       plot = lm_plot,
       width = 20,
       height = 12,
       units = "cm")

@@

```{r}
main %>% 
  filter(time_span == "preparation") %>% 
  group_by(ID) %>% 
  transmute(heart_rate_m = mean(heart_rate_std),
            confidence_factor_m = mean(confidence_factor,
                                      na.rm = TRUE),
            disruption_factor_m = mean(disruption_factor,
                                       na.rm = TRUE),
            teaching_experience = teaching_experience,
            gender = gender
  ) %>%
  distinct() %>% 
  ungroup() %>% 
  transmute(heart_rate_m = (heart_rate_m - mean(.$heart_rate_m))/sd(.$heart_rate_m),
         confidence_factor_m = (confidence_factor_m - mean(.$confidence_factor_m))/sd(.$confidence_factor_m),
         disruption_factor_m = (disruption_factor_m - mean(.$disruption_factor_m))/sd(.$disruption_factor_m),
         # gender = if_else(gender == "male",
         #                   true = 0,
         #                   false = 1),
          teaching_experience = (teaching_experience - mean(.$teaching_experience))/sd(.$teaching_experience)
         ) -> cor_pre

cor.test(cor_pre$heart_rate_m, cor_pre$confidence_factor_m)

@@

```{r}
df_multi <- 
  main %>% 
  select("ID", "disruption_factor", "confidence_factor", "teaching_experience") %>% 
  distinct() %>% 
  group_by(ID) %>% 
  summarize(disruption = mean(disruption_factor, na.rm = T),
            confidence = mean(confidence_factor, na.rm = T),
            teaching_experience) %>% 
  ungroup() %>% 
  mutate(id = as.character(ID)) %>% 
  distinct() %>% 
  right_join(x = .,
            y = fix_models,
            by = "id") %>% 
  mutate(slope = abs(slope))
  # mutate(posneg = if_else(condition = slope < 0,
  #                         true = "negative",
  #                         false = "positive"))

@@

# left_join(x = df_heart_rate,
#           y = df_heart_rate_std,
#           by = c("ID","time")) %>% 
bind_rows(df_heart_rate,
          df_heart_rate_std
          ) %>% 
  ggplot(mapping = aes(x = time,
                       y = heart_rate
                       )
         ) +
  # geom_point(alpha = 0.05) + 
  geom_smooth(se = T,
              size = 0.5,
              colour = "black",
              level = 0.99) +
  scale_color_viridis_d(option = "A",
                        end = 0.8) +
  scale_linetype_manual(values = c(1, 6)) +
  # scale_x_continuous(expand = c(0, NA),
  #                    breaks = seq(0, 120, 10)) +
  scale_x_continuous(expand = c(0.01,0)) + 
  labs(x = "Time (in minutes)",
       #y = "Heart Rate (in Beats per Minute)",
       y = "",
       # title = "Figure #",
       # subtitle = "Overall Course of Average Heartrate",
#        caption = expression(paste(italic("Note: "),
# "The shadow around the line represents the 99% confidence interval. The confidence interval shown refers to the HR measurement points during the entire study period. We used the ggplot2 package (v3.3.3; Wickham, 2016) to calculate the moving average of the course."))
       ) +
  #ggtitle("Overall Course of Average Heartrate") +
  theme_apa() + 
  # theme(plot.subtitle = element_text(face = "italic"),
  #       plot.caption = element_text(hjust = 0),
  #       axis.line = element_line(colour = "black"),
  #       panel.grid.major = element_blank(),
  #       panel.grid.minor = element_blank(),
  #       panel.border = element_blank(),
  #       panel.background = element_blank()) +
  facet_wrap(~ type,
             ncol = 2,
             scales = "free_y")

@@

```{r}

main %>% 
  filter(time_span != "overall") %>%
  dplyr::select("ID","time_span", "time","heart_rate_std") %>% 
  mutate(time_span = factor(time_span,
                            levels = c("preparation","teaching","post","interview","end"),
                            labels = c("(1) Pre-teaching\nInterval","(2) Teaching\nInterval","(3) Post-teaching\nInterval","(4) Interview\nInterval","(5) End\nInterval"))) %>% 
  distinct() %>%
  group_by(time_span) %>% 
  summarise(z_mean = mean(heart_rate_std),
            z_sd = sd(heart_rate_std),
            z_t = qt(p = 0.01 / 2,      # 0.01 KI
               df = n() - 1,
               lower.tail = F
               ),
            z_error = z_sd / sqrt(n()),
            lower = z_mean - z_t * z_error,
            upper = z_mean + z_t * z_error,
            ) %>% 
  ungroup() %>% 
  ggplot(mapping = aes(x = factor(0),
                       y = z_mean)) + 
  geom_point() +
  geom_errorbar(mapping = aes(ymin = lower,
                              ymax = upper),
                width = 0.2) +
  geom_hline(yintercept = 0,
             linetype = "dashed") +
  ylim(-1, 1) + 
  scale_y_continuous(breaks=seq(-1, 1, 0.5), 
                     limits=c(-1, 1)) + 
  labs(x = "Time (in Minutes)",
       y = "Standardized Mean Heart Rate",
       # caption = expression(paste(italic("Note: "),
       #                            "The shadow around the line represents the 99% confidence interval."
       #                            )
       #                      )
       ) +
  facet_grid(cols = vars(time_span),
             scales = "free_x") +
  theme_apa() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.caption = element_text(hjust = 0))

@@

```{r echo = FALSE, results='asis'}
df_multi <- 
  main %>% 
  select("ID", "disruption_factor", "confidence_factor", "teaching_experience") %>% 
  distinct() %>% 
  group_by(ID) %>% 
  summarize(disruption = mean(disruption_factor, na.rm = T),
            confidence = mean(confidence_factor, na.rm = T),
            teaching_experience) %>% 
  ungroup() %>% 
  mutate(id = as.character(ID)) %>% 
  distinct() %>% 
  right_join(x = .,
            y = fix_models,
            by = "id") %>%
  # mutate(posneg = if_else(condition = slope < 0,
  #                         true = "negative",
  #                         false = "positive"))

@@

```{r echo = FALSE, results='asis'}
df_multi <- 
  main %>% 
  select("ID", "disruption_factor", "confidence_factor", "teaching_experience") %>% 
  distinct() %>% 
  group_by(ID) %>% 
  summarize(disruption = mean(disruption_factor, na.rm = T),
            confidence = mean(confidence_factor, na.rm = T),
            teaching_experience) %>% 
  ungroup() %>% 
  mutate(id = as.character(ID)) %>% 
  distinct() %>% 
  right_join(x = .,
            y = fix_models,
            by = "id") %>% 
  # mutate(posneg = if_else(condition = slope < 0,
  #                         true = "negative",
  #                         false = "positive"))
fit_teach_1 <-
  df_multi %>% 
  filter(Phase == "Teaching Phase") %>% 
  # split(.$posneg) %>% 
  map(~ lm(slope ~ disruption,
           data = .))
